---
title: "UN SDG"
author: "James Hicks"
date: 'October 2020'
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 3
    toc_float: yes
---
# Introduction

This report was prepared by: 

* James Hicks

This report was finalized on October 30th 2020.
This report is generated from an R Markdown file that includes all the R code necessary to produce the results described and embedded in the report.  

Executing of this R notebook requires some subset of the following packages:

* `LiblineaR`
* `RColorBrewer`
* `SnowballC`
* `caTools`
* `car`
* `caret`
* `class`
* `countrycode`
* `corrplot`
* `dplyr`
* `dplot`
* `factoextra`
* `ggmap`
* `ggplot2`
* `kableExtra`
* `knitr`
* `lubridate`
* `magrittr`
* `plotly`
* `randomForest`
* `readr`
* `reshape2`
* `rgeos`
* `rnaturalearth`
* `rnaturalearthdata`
* `stringr`
* `tidyr`
* `tidyverse`
* `tm`

These will be installed and loaded as necessary (code suppressed). 


```{r,include=FALSE,echo=FALSE}
library(rnaturalearth)
library(rnaturalearthdata)
library(ggplot2)
library(factoextra)
library(dplyr)
library(ggpubr)
#library(car) 
library(randomForest)
library(lubridate)
library(class)
library(caret)
#library(LiblineaR)
library(plotly)
library(ggplot2)
#library(ggmap)
library(readr)
library(magrittr)
library(dplyr)
library(reshape2)
library(stringr)
library(tidyr)
library(magrittr)
library(testthat)
library(plyr)
library(rworldmap)
#library(assertive,warn.conflicts=FALSE)
library(countrycode)
```

```{r,include=FALSE,echo=FALSE}
#install.packages("ggsignif",type="win.binary")
#install.packages("ggpubr",type="win.binary")
#library(ggpubr)
```



```{r,include=FALSE,echo=FALSE}
if (!require("countrycode")) {
   install.packages("countrycode")
   library(countrycode)
}
if (!require("rworldmap")) {
   install.packages("rworldmap")
   library(rworldmap)
}
if (!require("rgeos")) {
   install.packages("rgeos")
   library(rgeos)

}
if (!require("rnaturalearth")) {
   install.packages("rnaturalearth")
   library(rnaturalearth)
}
if (!require("rnaturalearthdata")) {
   install.packages("rnaturalearthdata")
   library(rnaturalearthdata)
}
if (!require("e1071")) {
   install.packages("e1071")
   library(e1071)
}
if (!require("ggpubr")) {
   install.packages("ggpubr")
   library(ggpubr)
}

if (!require("factoextra")) {
   install.packages("factoextra")
   library(factoextra)
}
if (!require("ggplot2")) {
   install.packages("ggplot2")
   library(ggplot2)
}
if (!require("tidyverse")) {
   install.packages("tidyverse")
   library(tidyverse)
}
if (!require("readr")) {
   install.packages("readr")
   library(readr)
}
if (!require("dplyr")) {
   install.packages("dplyr")
   library(dplyr)
}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("corrplot")) {
   install.packages("corrplot")
   library(corrplot)
}
if (!require("factoextra")) {
   install.packages("factoextra")
   library(factoextra)
}
if (!require("kableExtra")) {
   install.packages("kableExtra")
   library(kableExtra)
}
if (!require("caTools")) {
   install.packages("caTools")
   library(caTools)
}
if (!require("randomForest")) {
   install.packages("randomForest")
   library(randomForest)
}
if (!require("tm")) {
   install.packages("tm")
   library(tm)
}
if (!require("SnowballC")) {
   install.packages("SnowballC")
   library(SnowballC)
}
if (!require("RColorBrewer")) {
   install.packages("RColorBrewer")
   library(RColorBrewer)
}

```


# Reading and Cleaning

In this section the data is read in and cleaned before data exploration and analysis.

## Reading
The chunk below reads in the data

```{r}
SDG_Chores<-read.csv("C:/Users/JHicks/Documents/R/UN_SDG/SDG_Chores.csv")
UN_Divorce<-read.csv("C:/Users/JHicks/Documents/R/UN_SDG/Un_Crude_Divorce_csv.csv")
```

## Cleaning

The chunk below renames the first colunm to "id" then removes redundant dimensons. Most of the colunms removed have a single value for the entire colunm.

```{r}
colnames(SDG_Chores)[1]<-"id"
SDG_Chores_ID<-SDG_Chores$id
#The following names are dropped due to redundancy. Most of them are static values
drop_names<-c("goal_code","goal_labelEN","goal_descEN","target_code","target_descEN","indicator_code","indicator_reference","indicator_descEN","series_release","series_tags","series","type","seriesDescription","Sources","footnotes","valueDetails","geoInfoUrl","timeDetails","Country_Profile","id")

SDG_Chores<-SDG_Chores[,!(names(SDG_Chores) %in%drop_names)]

```





The chunk below isolates and type casts non numeric data to a numeric format.

```{r}
num<-unlist(lapply(SDG_Chores,is.numeric))#isolate numeric values
SDG_Chores_Numeric<-SDG_Chores[,num]

#add variables that can be used in model construction or analysis, and convert them to numeric values if necessary
SDG_Chores_Numeric[["age_code"]]<-SDG_Chores$age_code
SDG_Chores_Numeric[["parentCode"]]<-SDG_Chores$parentCode
SDG_Chores_Numeric[["age_code"]]<-as.numeric(SDG_Chores_Numeric$age_code)
SDG_Chores_Numeric[["location_desc"]]<-as.numeric(SDG_Chores$location_desc)
SDG_Chores_Numeric[["sex_code"]]<-as.numeric(SDG_Chores$sex_code)

```


The chunk below removes the individual values for each year and stores them in a separate data set. This was done because the dataset is sparse. It is infeasible to compare data from 2000 to data from 2018, as such the most recent or latest value for each country was used.

```{r}

# remove the numeric year values and store them for future use. These colunms are too sparse for usage in PCA/RF
year_names<-c("value_2000","value_2001","value_2002","value_2003","value_2004","value_2005","value_2006","value_2007","value_2008","value_2009","value_2010","value_2011","value_2012","value_2013","value_2014","value_2015","value_2016","value_2017","value_2018")
SDG_Chores_Numeric_no_year<-SDG_Chores_Numeric[,!(names(SDG_Chores_Numeric) %in%year_names)]
SDG_Chores_Year_Values<-SDG_Chores_Numeric<-SDG_Chores_Numeric[,(names(SDG_Chores_Numeric) %in%year_names)]
```





# PCA Clustering

In this section, principal component analysis is performed on the numerical data to explain the variance in the data set. The prinipal components are then clustered and plotted in an attempt to understand the clustering present in the dataset. 

## PCA

The chunk below calculates the principal components and displays what proportion of variance is explained by each component.

```{r}
SDG_Chore_PCA<-prcomp(SDG_Chores_Numeric_no_year)
summary(SDG_Chore_PCA)#PC1+2 is ~94.4%
```

The chunk below displays the magnitude of the eigenvectors which convey the variables of greatest impact for each principal component. 

```{r}
SDG_Chore_PCA$rotation[,1:2]#botttom 3 are highest impact
```

### PCA & Latest Percentage

```{r}
#all of the samples projected onto PC1,PC2 colored by development (urban/rural/both)
colorScale<-brewer.pal(5,'RdPu')
Legend<-SDG_Chores$latest_value
p1<-ggplot(SDG_Chores_Numeric_no_year,aes(x=SDG_Chore_PCA$x[,1],y=SDG_Chore_PCA$x[,2],colour=Legend))+geom_point(size=2)
p1+ ggtitle("UN Subpopulations Denoted by Latest Percentage of Time Spent") +
  xlab("PC1") + ylab("PC2")+scale_color_gradientn(colours = rainbow(4))
```


### PCA & Age

```{r}
#all of the samples projected onto PC1,PC2 colored by development (urban/rural/both)
Legend<-as.factor(SDG_Chores$age_desc)
p1<-ggplot(SDG_Chores_Numeric_no_year,aes(x=SDG_Chore_PCA$x[,1],y=SDG_Chore_PCA$x[,2],colour=Legend))+geom_point(size=2)
p1+ ggtitle("UN Subpopulations Denoted by Reported Age Categories") +
  xlab("PC1") + ylab("PC2")
```



### PCA & Biological Sex

The chunk below plots the samples onto space spanned by the first two principal components, and colors them by biological sex.

```{r}
#all of the samples projected onto PC1,PC2 colored by the sex designation
Legend<-as.factor(SDG_Chores$sex_desc)
p1<-ggplot(SDG_Chores_Numeric_no_year,aes(x=SDG_Chore_PCA$x[,1],y=SDG_Chore_PCA$x[,2],colour=Legend))+geom_point(size=2)
p1+ ggtitle("UN Subpopulations Denoted by Biological Sex") +
  xlab("PC1") + ylab("PC2")
```


### PCA & Location Description

The chunk below plots the samples onto space spanned by the first two principal components, and colors them by how urban or rural the location is.

```{r}
#all of the samples projected onto PC1,PC2 colored by development (urban/rural/both)
Legend<-as.factor(SDG_Chores$location_desc)
p1<-ggplot(SDG_Chores_Numeric_no_year,aes(x=SDG_Chore_PCA$x[,1],y=SDG_Chore_PCA$x[,2],colour=Legend))+geom_point(size=2)
p1+ ggtitle("UN Subpopulations Denoted by Location Description") +
  xlab("PC1") + ylab("PC2")
```

### PCA & Biological Sex

The chunk below renames the various locations. This is done to enable better clustering. The colors in the plots are assigned alphabetically, so changing "Central America", "South America", and "North America" to "America Central", "America South", "America North" would color all of America similarly.

```{r}
parent_name_original<-SDG_Chores$parentName
parent_name_original_levels<- levels(parent_name_original)
parent_name_new_levels<-parent_name_original_levels
parent_name_new_levels[3]<-"America Central"
parent_name_new_levels[4]<-"Asia Central"
parent_name_new_levels[5]<-"Africa East"
parent_name_new_levels[6]<-"Asia East"
parent_name_new_levels[7]<-"Europe East"
parent_name_new_levels[9]<-"Africa Middle"
parent_name_new_levels[10]<-"Africa North"
parent_name_new_levels[11]<-"America North"
parent_name_new_levels[12]<-"Europe North"
parent_name_new_levels[13]<-"Asia South East"
parent_name_new_levels[14]<-"America South"
parent_name_new_levels[15]<-"Africa South"
parent_name_new_levels[16]<-"Asia South"
parent_name_new_levels[17]<-"Europe South"
parent_name_new_levels[18]<-"Africa West"
parent_name_new_levels[19]<-"Asia West"
parent_name_new_levels[20]<-"Europe West"

SDG_Chores$parentName<-mapvalues(SDG_Chores$parentName,from = parent_name_original_levels, to = parent_name_new_levels)
unsorted_levels<-levels(SDG_Chores$parentName)
SDG_Chores$parentName<-factor(SDG_Chores$parentName, levels=c(sort(unsorted_levels, decreasing=T)))
```



The chunk below plots the samples onto space spanned by the first two principal components, and colors them by the geographical parent area.

```{r}
#all of the samples projected onto PC1,PC2 colored by the parent region
Legend<-as.factor(SDG_Chores$parentName)
p1<-ggplot(SDG_Chores_Numeric_no_year,aes(x=SDG_Chore_PCA$x[,1],y=SDG_Chore_PCA$x[,2],colour=Legend))+geom_point(size=2)
p1+ ggtitle("UN Subpopulations Denoted by Parent Region") +
  xlab("PC1") + ylab("PC2")
```




## K-means & PCA

In this section the principal components are clustered using the k-means algorithm. There are not enough principal components to pursue more sophisicated clustering methods.

### 2 Clusters

The chunk below executes the k-means algorithm with 2 clusters and plots the variables of highest magnitude, colored by their cluster.

```{r}
set.seed(6432)
#k - means on the principal components with 2 clusters
var<-get_pca_var(SDG_Chore_PCA)
vertical_km<-kmeans(var$coord,centers=2,nstart=50)#k means with 2 clusters, 50 starts
clust<-as.factor(vertical_km$cluster)
fviz_pca_var(SDG_Chore_PCA,col.var=clust,palette=c("red", "blue", "green"),legend.title="Legend",repel=TRUE,title="K Means of PCA Variables - 2 Clusters")
```

### 3 Clusters

The chunk below executes the k-means algorithm with 3 clusters and plots the variables of highest magnitude, colored by their cluster.

```{r}
#k - means on the principal components with 3 clusters
vertical_km<-kmeans(var$coord,centers=3,nstart=50)
clust<-as.factor(vertical_km$cluster)
fviz_pca_var(SDG_Chore_PCA,col.var=clust,palette=c("red", "blue", "green"),legend.title="Legend",repel=TRUE,title="K Means of PCA Variables - 3 Clusters")
```

# Modeling

In this section, the data is modeled in attempt to classify the percentage of time spent, and again to classify the geographical parent area.

## Model Data Preparation

In this section a response variable is created that reflects 10% per category. A smaller margin could be used however there is not enough data to pursue more categories. There are no observed samples over 50% of time spent.

```{r}
#converting the latest percentage into categorical variables for easier clustering. 
#these values were chosen based on the distribution. there are only 9 samples above 30%
# 1 = 0-5.9999%
# 2 = 6-11.999% etc
response_5<-SDG_Chores_Numeric_no_year$latest_value
response_5_new<-as.character(SDG_Chores_Numeric_no_year$latest_value)
for (i in 1:length(response_5)){
  if (response_5[i]>=0 &response_5[i]<6)
    response_5_new[i]<-as.character("A")
  if (response_5[i]>=6 &response_5[i]<12)
    response_5_new[i]<-as.character("B")
  if (response_5[i]>=12 &response_5[i]<18)
    response_5_new[i]<-as.character("C")
  if (response_5[i]>=18 &response_5[i]<24)
    response_5_new[i]<-as.character("D")
  if (response_5[i]>=24 &response_5[i]<100)
    response_5_new[i]<-as.character("E")
}
SDG_Chores_Numeric_no_year[["response_5"]]<-as.factor(response_5_new)
Latest_Value_Copy<-SDG_Chores_Numeric_no_year$latest_value#copy the latest values for future usage
SDG_Chores_Numeric_no_year<-SDG_Chores_Numeric_no_year[, -which(names(SDG_Chores_Numeric_no_year) %in% c("latest_value"))]

```

The chunk below isolates the training and testing data for model construction.

```{r}
#isolating data for the random forest / NN
SDG_Chores_Numeric_no_year[["id"]]<-SDG_Chores_ID
ss<-as.integer(.8*nrow(SDG_Chores_Numeric_no_year))# 80% training / 20% testing
value<-sample(1:nrow(SDG_Chores_Numeric_no_year),ss)
traindata<-SDG_Chores_Numeric_no_year[value,]
testdata<-SDG_Chores_Numeric_no_year[-value,]
trainID<-traindata$id
testID<-testdata$id
traindata<-traindata[, -which(names(traindata) %in% c("id"))]
testdata<-testdata[, -which(names(testdata) %in% c("id"))]
```




## Random Forest

This section contains all random forest models.

### Latest Percentage Response

The chunk below creates a random forest from the training data.

#### Random Forest Training
```{r,echo=FALSE,results=FALSE}
attach(traindata)
randFor<-randomForest::randomForest(response_5~.,data=traindata)
```

The chunk below prints the details of the random forest. 

```{r}
print(randFor)
```

The chunk below prints the change in the cumulative error rate as more decision trees are generated. 

```{r}

layout(matrix(c(1,2),nrow=1),width=c(4,1)) 
plot(randFor,main="Random Forest Error: Latest Value Percentage Category")
par(mar=c(4,0,4,1)) #L
plot(c(.5,1),type="n", axes=F, xlab="", ylab="")
legend("top",colnames(randFor$err.rate),col=1:5,cex=.9,fill = 1:5)
```

The chunk below displays the top variables used to improve the model performance. Note the custom variable names. if any changes are made to the random forest model these labels may not be valid anymore.

```{r}
#,labels=rep("random variable name",5)
randomForest::varImpPlot(randFor,sort=T,n.var=10,main="Variable Importance - Classifying Latest Value Percentage",labels=rep(c("Years of Data","Location Description","Maximum Year","Parent Region","Minimum Year","Geographic Area","X","Y","Age Category","Sex")))
```
#### Random Forest Testing

The chunk below prints the percentage correct in testing for the random forest model.

```{r}
fortable<-table(predict(randFor,testdata),as.factor(testdata$response_5))
tmp<-diag(prop.table(fortable,1))
#printing testing statistics
cat("\nError Rate in Testing\n\tCategory 1:\t",100-round(tmp[1]*100,2),"\n\tCategory 2:\t",100-round(tmp[2]*100,2),"\n\tCategory 3:\t",100-round(tmp[3]*100,2),"\n\tCategory 4:\t",100-round(tmp[4]*100,2),"\n\tCategory 5:\t",100-round(tmp[5]*100,2),"\n\tAverage:\t",100-round(sum(diag(prop.table(fortable)))*100,2))
```

The chunk below prints the confusion matrix for the testing set of the random forest model.

```{r}
cat("\n","\nConfusion Matrix:\n")
fortable
```

### Parent Code Response

In this section, the same workflow is followed with the exception of a different response variable. This section's response variable represents the geographical parent region name.


```{r}
#response variable of the parent region name instead of the % of work
Latest_Value_Copy<-SDG_Chores_Numeric_no_year$latest_value
SDG_Chores_Parent_RF<-SDG_Chores_Numeric_no_year
SDG_Chores_Parent_RF[["latest_value"]]<-Latest_Value_Copy
SDG_Chores_Parent_RF[["response_region"]]<-as.factor(SDG_Chores$parentName)
SDG_Chores_Parent_RF<-SDG_Chores_Parent_RF[, -which(names(SDG_Chores_Parent_RF) %in% c("geoAreaCode","parentCode","response_5","parentName","X","Y","location_desc"))]
```

```{r}
ss2<-as.integer(.8*nrow(SDG_Chores_Parent_RF))# 80 / 20 
value2<-sample(1:nrow(SDG_Chores_Parent_RF),ss2)
traindata2<-SDG_Chores_Parent_RF[value2,]
testdata2<-SDG_Chores_Parent_RF[-value2,]
trainID2<-traindata2$id
testID2<-testdata2$id
traindata2<-traindata2[, -which(names(traindata2) %in% c("id"))]
testdata2<-testdata2[, -which(names(testdata2) %in% c("id"))]
```

#### RF Training
#### Random Forest Training


```{r,echo=FALSE,results=FALSE}

attach(traindata2)
randFor2<-randomForest::randomForest(response_region~.,data=traindata2)
```

```{r}
print(randFor2)
```

```{r}
#layout(matrix(c(1,2),nrow=1),width=c(4,1)) 
plot(randFor2,main="Random Forest Error: Region Category")
#par(mar=c(4,0,4,1)) #L
#plot(c(.5,1),type="n", axes=F, xlab="", ylab="")
#legend("top",colnames(randFor2$err.rate),col=1:5,cex=.9,fill = 1:5)
#there are too many categories for a label to be done quickly here - would require remaking the plot with gg
```

```{r}
randomForest::varImpPlot(randFor2,sort=T,n.var=7,main="Variable Importance - Classifying Region",labels=rep(c(
  "Sex","UN Member Status","Age Category","Years of Data","level_","Maximum Year","Minimum Year")))
```


#### RF Testing
The chunk below displays the percentage correct in testing for each category. Unfortunately there is not enough data for certain areas such as the Caribbean, which results in no observed value for certain categories.

```{r}
fortable2<-table(predict(randFor2,testdata2),as.factor(testdata2$response_region))
table2_names<-row.names(fortable2)
tmp<-diag(prop.table(fortable2,1))


cat("\nError Rate in Testing\n\t",table2_names[1],":\t",100-round(tmp[1]*100,2),"\n\t",table2_names[2],":\t",100-round(tmp[2]*100,2),"\n\t",table2_names[3],":\t",100-round(tmp[3]*100,2),"\n\t",table2_names[4],":\t",100-round(tmp[4]*100,2),"\n\t",table2_names[5],":\t",100-round(tmp[5]*100,2),"\n\t ",table2_names[6],":\t",100-round(tmp[6]*100,2),"\n\t",table2_names[7],":\t",100-round(tmp[7]*100,2),"\n\t",table2_names[8],":\t",100-round(tmp[8]*100,2),"\n\t",table2_names[9],":\t",100-round(tmp[9]*100,2),"\n\t",table2_names[10],":\t",100-round(tmp[10]*100,2),"\n\t",table2_names[11],":\t",100-round(tmp[11]*100,2),"\n\t",table2_names[12],":\t",100-round(tmp[12]*100,2),"\n\t",table2_names[13],":\t",100-round(tmp[13]*100,2),"\n\t",table2_names[14],":\t",100-round(tmp[14]*100,2),"\n\t",table2_names[15],":\t",100-round(tmp[15]*100,2),"\n\t",table2_names[16],":\t",100-round(tmp[16]*100,2),"\n\t",table2_names[17],":\t",100-round(tmp[17]*100,2),"\n\t",table2_names[18],":\t",100-round(tmp[18]*100,2),"\n\t",table2_names[19],":\t",100-round(tmp[19]*100,2),"\n\t",table2_names[20],":\t",100-round(tmp[20]*100,2),"\n\tAverage:\t",100-round(sum(diag(prop.table(fortable)))*100,2))
```



## Neural Network

### Latest Percentage Response

```{r,echo=FALSE,results=FALSE}
trainc<-trainControl(method="repeatedcv",number=10,repeats=3)
NN_5<-train(response_5~.,data=traindata,method="nnet",tuneLength=10,trainControl=trainc)
```

```{r}
test_NN_5<-predict(NN_5,newdata=testdata)#testing
NN_5_table<-table(test_NN_5,testdata$response_5)
NN_5_table
diag(prop.table(NN_5_table,1))# % correct in each class IN test
cat("Percentage Correct in Testing: ",sum(diag(prop.table(NN_5_table))))#overalll % correct IN Test
```



### Parent Code Response

```{r,echo=FALSE,results=FALSE,message=FALSE,}
NN_region<-train(response_region~.,data=traindata2,method="nnet",tuneLength=10,trainControl=trainc)
```

```{r}
test_NN_region<-predict(NN_region,newdata=testdata2)#testing
NN_region_table<-table(test_NN_region,testdata2$response_region)
#NN_region_table
#diag(prop.table(NN_region_table,1))# % correct in each class IN test
cat("Percentage Correct in Testing: ",sum(diag(prop.table(NN_region_table))))#overalll % correct IN Test
```




# World Plots with Chore Percentage


## Data Isolation
```{r}
plot_colunms<-c("ISO3","location_desc","latest_value","sex_code")
country_chores<-SDG_Chores[,(names(SDG_Chores) %in%plot_colunms)]
```

```{r}
earth<-ne_countries(scale="medium",returnclass ="sf")
```


```{r}
country_chores_rural_m<-country_chores%>%filter(location_desc=="Rural",sex_code=="M")
country_chores_rural_m<-country_chores_rural_m[,-3]#drop the sex colunm as the values are static
country_chores_rural_m<-aggregate(country_chores_rural_m[,3],list(country_chores_rural_m$ISO3),mean)#mean because of the different age groups

country_chores_rural_f<-country_chores%>%filter(location_desc=="Rural",sex_code=="F")
country_chores_rural_f<-country_chores_rural_f[,-3]
country_chores_rural_f<-aggregate(country_chores_rural_f[,3],list(country_chores_rural_f$ISO3),mean)
```

```{r}
country_chores_urban_m<-country_chores%>%filter(location_desc=="Urban",sex_code=="M")
country_chores_urban_m<-country_chores_urban_m[,-3]#drop the sex colunm as the values are static
country_chores_urban_m<-aggregate(country_chores_urban_m[,3],list(country_chores_urban_m$ISO3),mean)

country_chores_urban_f<-country_chores%>%filter(location_desc=="Urban",sex_code=="F")
country_chores_urban_f<-country_chores_urban_f[,-3]
country_chores_urban_f<-aggregate(country_chores_urban_f[,3],list(country_chores_urban_f$ISO3),mean)
```

```{r}
country_chores_all_m<-country_chores%>%filter(location_desc=="All areas",sex_code=="M")
country_chores_all_m<-country_chores_all_m[,-3]#drop the sex colunm as the values are static
country_chores_all_m<-aggregate(country_chores_all_m[,3],list(country_chores_all_m$ISO3),mean)

country_chores_all_f<-country_chores%>%filter(location_desc=="All areas",sex_code=="F")
country_chores_all_f<-country_chores_all_f[,-3]
country_chores_all_f<-aggregate(country_chores_all_f[,3],list(country_chores_all_f$ISO3),mean)
```


```{r,results=FALSE,echo=FALSE}
chores_rural_m_map<-joinCountryData2Map(country_chores_rural_m, joinCode="ISO3", nameJoinColumn="Group.1")
chores_rural_f_map<-joinCountryData2Map(country_chores_rural_f, joinCode="ISO3", nameJoinColumn="Group.1")

chores_urban_m_map<-joinCountryData2Map(country_chores_urban_m, joinCode="ISO3", nameJoinColumn="Group.1")
chores_urban_f_map<-joinCountryData2Map(country_chores_urban_f, joinCode="ISO3", nameJoinColumn="Group.1")

chores_all_m_map<-joinCountryData2Map(country_chores_all_m, joinCode="ISO3", nameJoinColumn="Group.1")
chores_all_f_map<-joinCountryData2Map(country_chores_all_f, joinCode="ISO3", nameJoinColumn="Group.1")
```
## Statistics

```{r}
#statistics rural m


summary(chores_rural_m_map$x)
```

```{r}
#statistics Urban m

summary(chores_urban_m_map$x)
```



```{r}
#statistics Rural F

summary(chores_rural_f_map$x)
```


```{r}
#statistics Urban F

summary(chores_urban_f_map$x)
```

## World plots 


### Rural


```{r}
#rural M
colorScale_r<-brewer.pal(9,"Reds")
colorScale_u<-brewer.pal(11,"Blues")
map_variables_r_m<-mapCountryData(chores_rural_m_map,nameColumnToPlot ="x",oceanCol = "azure1",catMethod = "pretty",missingCountryCol = gray(.7),addLegend = T,mapTitle = "Percentage of Time spent on Domestic work - Rural Male Population", borderCol ="black",colourPalette = colorScale_r)

```


```{r}
#rural F


map_variables_r_m<-mapCountryData(chores_rural_f_map,nameColumnToPlot ="x",oceanCol = "azure1",catMethod = "pretty",missingCountryCol = gray(.7),addLegend = T,mapTitle = "Percentage of Time spent on Domestic work - Rural Female Population", borderCol ="black",colourPalette = colorScale_r)

```

### Urban

```{r}
#Urban M


map_variables_r_m<-mapCountryData(chores_urban_m_map,nameColumnToPlot ="x",oceanCol = "azure1",catMethod = "pretty",missingCountryCol = gray(.7),addLegend = T,mapTitle = "Percentage of Time spent on Domestic work - Urban Male Population", borderCol ="black",colourPalette = colorScale_u)

```

```{r}
#Urban f
map_variables_r_m<-mapCountryData(chores_urban_f_map,nameColumnToPlot ="x",oceanCol = "azure1",catMethod = "pretty",missingCountryCol = gray(.7),addLegend = T,mapTitle = "Percentage of Time spent on Domestic work - Urban Female Population", borderCol ="black",colourPalette = colorScale_u)

```


### All Areas

```{r}
#All M

colorScale_a<-brewer.pal(9,"Greens")
map_variables_r_m<-mapCountryData(chores_all_m_map,nameColumnToPlot ="x",oceanCol = "azure1",catMethod = "pretty",missingCountryCol = gray(.7),addLegend = T,mapTitle = "Percentage of Time spent on Domestic work - Male Population", borderCol ="black",colourPalette = colorScale_a)

```

```{r}
#All F


map_variables_r_m<-mapCountryData(chores_all_m_map,nameColumnToPlot ="x",oceanCol = "azure1",catMethod = "pretty",missingCountryCol = gray(.7),addLegend = T,mapTitle = "Percentage of Time spent on Domestic work - Female Population", borderCol ="black",colourPalette = colorScale_a)

```

# World Plots with Divorce/Separation Percentage



## Data Isolation

This chunk primarily removes the headers and data from prior to before the year 2000

```{r}
UN_Divorce_clean<-UN_Divorce[-1,]
colnames(UN_Divorce_clean)[1]<-"Country"
colnames(UN_Divorce_clean)[2]<-"ISO_Code"
colnames(UN_Divorce_clean)[15]<-"Latest_Year"
colnames(UN_Divorce_clean)[16]<-"CDR"
UN_Divorce_clean<-UN_Divorce_clean[-1,]#white space and colnum information
UN_Divorce_clean<-UN_Divorce_clean[-1,]
UN_Divorce_clean<-UN_Divorce_clean[-1,]
```

```{r,echo=FALSE,results=FALSE}
divorce_keep_colunms<-c("ISO_Code","CDR")
country_divorce_all<-UN_Divorce_clean[,names(UN_Divorce_clean)%in%divorce_keep_colunms]
country_divorce_all$CDR<-as.numeric(levels(country_divorce_all$CDR))[country_divorce_all$CDR]
country_divorce_all$ISO_Code<-countrycode(country_divorce_all$ISO_Code,origin='genc3n',destination='iso3c')
UN_Divorce_map_data<-joinCountryData2Map(country_divorce_all, joinCode="ISO3", nameJoinColumn="ISO_Code")
```


## World Plot

```{r}
colorScale_divorce<-brewer.pal(10,"YlGnBu")
map_variables_r_m<-mapCountryData(UN_Divorce_map_data,nameColumnToPlot ="CDR",oceanCol = "azure1",catMethod = "pretty",missingCountryCol = gray(.7),addLegend = T,mapTitle = "Worldwide Crude Divorce Rate - UN Statistics 2007", borderCol ="black",colourPalette = colorScale_divorce)
```


